Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-10-27 ms.tex
Fixing signs starting with Eq. 39 - integrals over t should have an extra -

Taking a derivative with respect to $r$ and evaluating at the surface $r=R$, we use
\be
\frac{dj_0(\kappa_n r)}{dr} \bigg\rvert_R & =& \frac{d}{dr} \left( \frac{\sin(\kappa_n r)}{\kappa_n r} \right)\bigg\rvert_R
=  \left( \frac{\cos(\kappa_n R)}{R} - \frac{\sin(\kappa_n R)}{\kappa_n R^2} \right)
\nonumber \\ & = & \frac{(-1)^n}{R}
\ee
to obtain the flux, which is
\be
F(R,\sigma,t) & =& - \frac{4\pi}{3k\phi} \frac{dJ(R,\sigma,t)}{dr} 
= - \frac{4\pi}{3k\phi R}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
% TODO: PICK UP HERE
Multiplying by $4\pi R^2$ gives the energy per time per frequency
\be
\frac{dE}{dtd\nu} & = & - \frac{16\pi^2 R}{3k\phi}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\label{eq:dEdtdnu}
\ee
Integrating over time yields a factor $-1/\gamma_{nm}$, and by integrating over $d\nu$ we find
\be \label{eq:sum_rule}
1 & = &  \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE} \sum_{nm} (-1)^n \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
This non-trivial ``sum rule'' provides a check on the values of $\gamma_{nm}$ and $J_{nm}(\sigma)$. This expression can also be written as
\be
1 & =& \sum_{nm} P_{nm},
\label{eq:sumrule}
\ee
where the contribution of each mode is
\be \label{eq:pnmsoln}
P_{nm} & \equiv & \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE}  (-1)^n \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
These coefficients are negative for odd values of $n$ and positive for even $n$. The size of each contribution scales roughly as $0.5/(m-7/8)^{2/3}$, with a weak dependence on $n$. This indicates the need for a large number of $n$ and $m$ to converge, in that it takes roughly ten times as many $m$ modes to reduce the size of $P_{nm}$ by a factor of ${\sim}$5.

@@

We now calculate the distribution of the times it takes each photon to escape the sphere. This is obtained by integrating Eq. \ref{eq:dEdtdnu} over all frequencies. We find
\be
P(t)  & = & \sqrt{\frac{3}{2}} \frac{16\pi^2 R \Delta^2 }{3kE}     \sum_{nm} (-1)^n  e^{-\gamma_{nm}t} \int d\sigma J_{nm}(\sigma) 
\nonumber \\ & = &  \sum_{nm} P_{nm} \gamma_{nm} e^{-\gamma_{nm}t},
\label{eq:waittime}
\ee
which is normalized to unity. For a sufficiently large number of spatial modes $n$ and frequency modes $m$, the result of this sum agrees with Monte Carlo escape time distributions. % Not sure if we want to say this yet? Maybe save for the escape time figure.

The eigenfunction expansion method utilized here can produce an effective characterization of the escape time distribution, yet it is not a fast enough computational method to use as a distribution for sampling. However, the behavior of the solution is well-understood since the late-time distribution is simply an exponential falloff. The rate constant of the falloff is the lowest-order eigenfrequency, $\gamma_{00}$, and its scale is determined by the coefficient $P_{00}$ as in Eq. \ref{eq:pnmsoln}. Thus, an approximate ``fitting function'' that captures both the peak of the escape time distribution and the exponential falloff is
\be \label{eq:fitting_function}
P(t) = \exp{\left[-\left(\frac{t_{\rm diff}}{t}\right)^2\right]} \times \gamma_{00} P_{00} e^{-\gamma_{00}t}.
\ee
The first term represents the early-time distribution, which then transitions to an exponential falloff past a point $ct/R = ct_{\rm diff}/R = (a\tau_0)^{1/3}$, where $t_{\rm diff}$ is a characteristic diffusion timescale.

@@

We now evaluate the time-integrated spectrum of the response to an impulse and compare it with the solution for the $H_0$ steady-state spectrum (Eq. \ref{eq:H0surf}). Integrating Eq. \ref{eq:dEdtdnu} over all times gives a distribution for the emitted frequencies. Dividing by the energy $E$, we find the probability distribution
\be \label{eq:spectrum}
P(\nu) & = &  \frac{16\pi^2 R}{3k\phi E}  \sum_{nm} (-1)^n \gamma_{nm}^{-1} J_{nm}(\sigma).
\ee
Integrating over $\nu$ then gives unity as required by the sum rule in Eq. \ref{eq:sumrule}.

>>>>

Taking a derivative with respect to $r$ and evaluating at the surface $r=R$, we use
\be
\frac{dj_0(\kappa_n r)}{dr} \bigg\rvert_R & =& \frac{d}{dr} \left( \frac{\sin(\kappa_n r)}{\kappa_n r} \right)\bigg\rvert_R
=  \left( \frac{\cos(\kappa_n R)}{R} - \frac{\sin(\kappa_n R)}{\kappa_n R^2} \right)
\nonumber \\ & = & \frac{(-1)^n}{R}
\ee
to obtain the flux, which is
\be
F(R,\sigma,t) & =& - \frac{4\pi}{3k\phi} \frac{dJ(R,\sigma,t)}{dr} 
= - \frac{4\pi}{3k\phi R}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Multiplying by $4\pi R^2$ gives the energy per time per frequency emerging from the sphere to be
\be
\frac{dE}{dtd\nu} & = & - \frac{16\pi^2 R}{3k\phi}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\label{eq:dEdtdnu}
\ee
Integrating over time yields a factor $1/\gamma_{nm}$, and by integrating over $d\nu$ we find
\be \label{eq:sum_rule}
1 & = &  \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE} \sum_{nm} (-1)^{n+1} \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
This non-trivial ``sum rule'' provides a check on the values of $\gamma_{nm}$ and $J_{nm}(\sigma)$. This expression can also be written as
\be
1 & =& \sum_{nm} P_{nm},
\label{eq:sumrule}
\ee
where the contribution of each mode is
\be \label{eq:pnmsoln}
P_{nm} & \equiv & \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE}  (-1)^{n+1} \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
These coefficients are negative for odd values of $n$ and positive for even $n$. The size of each contribution scales roughly as $0.5/(m-7/8)^{2/3}$, with a weak dependence on $n$. This indicates the need for a large number of $n$ and $m$ to converge, in that it takes roughly ten times as many $m$ modes to reduce the size of $P_{nm}$ by a factor of ${\sim}$5.

@@

We now calculate the distribution of the times it takes each photon to escape the sphere. This is obtained by integrating Eq. \ref{eq:dEdtdnu} over all frequencies. We find
\be
P(t)  & = & \sqrt{\frac{3}{2}} \frac{16\pi^2 R \Delta^2 }{3kE}     \sum_{nm} (-1)^n  e^{-\gamma_{nm}t} \int d\sigma J_{nm}(\sigma) 
\nonumber \\ & = &  \sum_{nm} P_{nm} \gamma_{nm} e^{-\gamma_{nm}t},
\label{eq:waittime}
\ee
which is normalized to unity. For a sufficiently large number of spatial modes $n$ and frequency modes $m$, the result of this sum agrees with Monte Carlo escape time distributions.

The eigenfunction expansion method utilized here can produce an effective characterization of the escape time distribution, yet it is not a fast enough computational method to use as a distribution for sampling. However, the behavior of the solution is well-understood since the late-time distribution is simply an exponential falloff. The rate constant of the falloff is the lowest-order eigenfrequency, $\gamma_{11}$, and its scale is determined by the coefficient $P_{11}$ as in Eq. \ref{eq:pnmsoln}. Thus, an approximate ``fitting function'' that captures both the peak of the escape time distribution and the exponential falloff is
\be \label{eq:fitting_function}
P(t) = \exp{\left[-\left(\frac{t_{\rm diff}}{t}\right)^2\right]} \times \gamma_{11} P_{11} e^{-\gamma_{11}t}.
\ee
The first term represents the early-time distribution, which then transitions to an exponential falloff past a point $ct/R = ct_{\rm diff}/R = (a\tau_0)^{1/3}$, where $t_{\rm diff}$ is a characteristic diffusion timescale.

@@

We now evaluate the time-integrated spectrum of the response to an impulse and compare it with the solution for the $H_0$ steady-state spectrum (Eq. \ref{eq:H0surf}). Integrating Eq. \ref{eq:dEdtdnu} over all times gives a distribution for the emitted frequencies. Dividing by the energy $E$, we find the probability distribution
\be \label{eq:spectrum}
P(\nu) & = &  \frac{16\pi^2 R}{3k\phi E}  \sum_{nm} (-1)^{n+1} \gamma_{nm}^{-1} J_{nm}(\sigma).
\ee
Integrating over $\nu$ then gives unity as required by the sum rule in Eq. \ref{eq:sumrule}.
